What paradigms can webcam eye-tracking be used for? Attempted replications of five ``classic'' cognitive science experiments

### 2024-09-01 manuscript.tex
Render Quarto Manuscript

\subsection{Data pre-processing}\label{data-pre-processing}

@@

\section{Experiment 1}\label{experiment-1}

@@

\paragraph{In-lab sample}\label{in-lab-sample}

Forty-nine participants completed the study in a lab setting. They were recruited via the college subject pool. Participants needed to be 18 years of age or older and native speakers of English. The study was approved by the Institutional Review Board at Boston College.

@@

\subsubsection{Materials \& Design}\label{materials-design}

@@

Each participant was presented with sixteen critical trials (eight in
the restrictive condition, eight in the non-restrictive condition) and
sixteen fillers for a total of 32 trials. The order of trials and the
assignment of critical scene to condition was random on a
subject-by-subject basis.

\subsubsection{Data pre-processing and analysis}\label{data-pre-processing-and-analysis}

@@

\subsubsection{Remote sample}\label{remote-sample-1}

@@

We ran a regression model predicting the timing of the first fixation to the target object, relative to the onset of the noun, with verb condition as a predictor, mean-centered verb duration as a covariate, and random intercepts and condition slopes for participants and scenes. There were no significant effects; participants looked sooner at the target in the restrictive condition, while accounting for verb duration and its interaction with condition, but this was not a statistically significant effect (\emph{b} = -121.91, \emph{SE} = 90.57, \emph{p}=0.20).

@@

@@

We tested whether a participant's calibration quality was correlated with their effect size. There were three effects of interest: the verb-by-object interaction in predicting fixation probabilities, both in the (1) pre-noun-onset and (2) pre-verb-offset windows (calculated as the difference in target-over-distractor preference between verb conditions), and (3) the effect of verb on the timing of the first target fixation (calculated as the difference in target latency between verb conditions). Across the three effects of interest, calibration quality was not significantly correlated (Effect 1: Pearson's \emph{r} = 0.03, \emph{p} = 0.83, Effect 2: Pearson's \emph{r} = -0.05, \emph{p} = 0.73, Effect 3: Pearson's \emph{r} = 0.04, \emph{p} = 0.78. In the remote sample, we saw that calibration quality was correlated with target advantage in the pre-noun window in the restricting condition alone. This was not the case in the in-lab sample (Pearson's \emph{r} = 0.21, \emph{p} = 0.14).

\subsubsection{In-lab sample}\label{in-lab-sample-1}

@@

As before, participants' calibration quality was measured as the mean percentage of fixations that landed within 200 pixels of the calibration point. Calibration quality ranged from 5.13\% to 97.89\%.

We tested whether a participant's calibration quality was correlated with their effect size. Across the three condition effects of interest, calibration quality was not significantly correlated (pre-noun: Pearson's \emph{r} = -0.24, \emph{p} = 0.11; pre-verb-offset: Pearson's \emph{r} = -0.19, \emph{p} = 0.20; first fixation: Pearson's \emph{r} = -0.12, \emph{p} = 0.41. However, when the two interaction effects are calculated as the target advantage in the restricting condition only (i.e.~rather than a difference of differences), we see a significant correlation between target advantage and calibration quality in the wider pre-noun window (Pearson's \emph{r} = -0.16, \emph{p} = 0.29).

@@

@@

@@

@@

@@

>>>>

\subsection{Data Pre-processing}\label{data-pre-processing}

@@

\section{Experiment 1a}\label{experiment-1a}

@@

@@

\subsubsection{Materials and Design}\label{materials-and-design}

@@

Each participant was presented with 16 critical trials (eight in
the restrictive condition, eight in the non-restrictive condition) and
16 fillers for a total of 32 trials. The order of trials and the
assignment of critical scene to condition was random on a
subject-by-subject basis.

@@

@@

We ran a regression model predicting the timing of the first fixation to the target object, relative to the onset of the noun, with verb condition as a predictor, mean-centered verb duration as a covariate, and random intercepts and condition slopes for participants and scenes\^{}{[} \texttt{lme4} syntax: \texttt{lmer\_alt(time\ \textasciitilde{}\ verb\_condition*verb\_duration\ +\ (verb\_condition\ \textbar{}\textbar{}\ subject)\ +\ (verb\_condition\ \textbar{}\textbar{}\ scene)}. There were no significant effects; participants looked sooner at the target in the restrictive condition, while accounting for verb duration and its interaction with condition, but this was not a statistically significant effect (\emph{b} = -121.91, \emph{SE} = 90.57, \emph{p}=0.20).

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/E1-calib-effect-1.pdf}
\caption{\label{fig:E1-calib-effect}Calibration scores plotted against target advantage scores (cumulative proportion of fixations to the target minus cumulative proportion of fixations to competitors) at the end of the pre-noun time window.}
\end{figure}

@@

We tested whether a participant's calibration quality was correlated with their effect size. There were three effects of interest: the verb-by-object interaction in predicting fixation probabilities, both in the (1) pre-noun-onset and (2) pre-verb-offset windows (calculated as the difference in target-over-distractor preference between verb conditions), and (3) the effect of verb on the timing of the first target fixation (calculated as the difference in target latency between verb conditions). Across the three effects of interest, calibration quality was not significantly correlated (Effect 1: Pearson's \emph{r} = 0.03, \emph{p} = 0.83, Effect 2: Pearson's \emph{r} = -0.05, \emph{p} = 0.73, Effect 3: Pearson's \emph{r} = 0.04, \emph{p} = 0.78. However, when the two interaction effects were calculated as the target advantage in the restricting condition only (i.e.~rather than a difference of differences), we see a significant correlation between target advantage and calibration quality in the wider pre-noun window (Pearson's \emph{r} = 0.21, \emph{p} = 0.14). This is shown in Figure~\ref{fig:E1-calib-effect}.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/E1a-spaghetti-fig-1.pdf}
\caption{\label{fig:E1a-spaghetti-fig}Cumulative probability of fixating distractor and target objects across conditions over time, with 0 ms aligned to the verb onset time. The vertical line marks the mean noun onset time across trials and conditions.}
\end{figure}

Across three different tests of the hypothesis that listeners will use verb semantics to anticipate the upcoming referent, we found results that were numerically consistent with the hypothesis but not statistically significant. A comparison to published data (James et al., 2023) demonstrates that looks to the objects in the scene (relative to background or off-screen looks) were depressed across conditions and objects. After eliminating participants with validation accuracy under 50\% and/or 10\% or fewer fixations to any ROIs, we were left with only 22 of the original 60 participants. Analyses on this subset resulted in one effect reaching statistical significance (shorter target fixation latencies in the restricting vs.~non-restricting verb condition).
Given that nearly two-thirds of participants had poor data quality, we ran a follow-up webcam study in a lab setting order to isolate the potential causes.

\section{Experiment 1b}\label{experiment-1b}

Experiment 1b tested whether the failure to replicate significant condition effects in Experiment 1a was due to features of conducting the study remotely (i.e.~varied experimental settings and apparatuses, lower compliance) rather than webcam-based eye-tracking or \texttt{webgazer} \emph{per se}. Thus, Experiment 1b took place in a lab setting with undergraduate participants but otherwise used the same Method as Experiment 1a.

Forty-nine participants completed the study in a lab setting. They were recruited via the Boston College subject pool. Participants needed to be 18 years of age or older and native speakers of English.

After being greeted by the experimenter and completing the informed consent form, participants followed on-screen prompts to complete the study, including calibration, as described in the Experiment 1a Procedure.

\subsubsection{Apparatus}\label{apparatus}

HELP

\subsubsection{Materials and Design}\label{materials-and-design-1}

Materials were identical to those in Experiment 1a.

@@

As before, participants' calibration quality was measured as the mean percentage of fixations that landed within 200 pixels of the calibration point. Calibration quality ranged from 5.13\% to 97.89\%.
Across the three condition effects of interest, calibration quality was not significantly correlated (pre-noun: Pearson's \emph{r} = -0.24, \emph{p} = 0.11; pre-verb-offset: Pearson's \emph{r} = -0.19, \emph{p} = 0.20; first fixation: Pearson's \emph{r} = -0.12, \emph{p} = 0.41.

In the remote sample, we saw that calibration quality was correlated with target advantage in the pre-noun window in the restricting condition alone. This was not the case in the in-lab sample (Pearson's \emph{r} = -0.16, \emph{p} = 0.29).

@@

\subsubsection{Procedure}\label{procedure-4}

@@

\subsection{Method}\label{method-5}

@@

\subsubsection{Participants}\label{participants-6}

@@

\subsection{Results}\label{results-5}

@@

\subsection{Discussion}\label{discussion-5}
